A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-10 2022/OM&S/paper.tex
Section 4

For optimization of convolutions a wide spectrum of algorithms can be used in dependence on convolution properties generated by the features of criteria in the initial MCO problem. In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach. Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.

@@

If number of completed trials is not exactly divisible by $J$, where \textit{global-to-local ratio} $J$ is a predefined number, then
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

@@

@@

>>>>

For optimization of convolutions a wide spectrum of algorithms can be used in dependence on convolution properties generated by the features of criteria in the initial MCO problem. In particular, for the multiextremal case where it is required to find the global optimum the publications \cite{Evtushenko2014,Zilinskas2015,GERGEL2017_1,Gergel2019_2,Barkalov2021} describe applications of the global optimization algorithms to MCO problems in combination with the convolution approach. Among these methods of multiextremal optimization the efficient algorithms that utilize  schemes of dimensionality reduction such as Peano mappings \cite{Strongin2000,Sergeyev2013} or nested optimization \cite{Strongin2000,Grishagin2015_2,Grishagin2018} are perspective for application in the  frameworks of multiobjective optimization.

@@

If number of completed trials is not exactly divisible by $T$, where \textit{global-to-local ratio} $T$ is a predefined number, then
\begin{equation}\label{eq:16}
R(q) = B(q),
\end{equation}
otherwise
\begin{equation}\label{eq:17}
R(q) = B(q)\left( \frac {\sqrt{(v_q - \varphi^*(n))(v_{q-1} - \varphi^*(n))}}{M} +(1.5)^{-\sigma}\right)^{-1}.
\end{equation}
Here
\begin{gather}\label{eq:18}
B(q) = (t_s - t_{s-1})^{1/N} + \frac{(v_q - v_{q-1})^2} {m^2 (t_s - t_{s-1})^{1/N}} -2\frac{v_{q-1} + v_q}{m},\\
\varphi^*(n) = \min_{1\leq j \leq n} {\varphi(t^j)}
\end{gather}
is minimal computed value of the objective function $\varphi(t)$ at trial points,
\begin{equation}\label{eq:19}
M = \max_{2 \leq  s \leq n} { \frac{|v_s - v_{s-1}|}{(t_s - t_{s-1})^{1/N}} }
\end{equation}
is underestimation of the H{\"o}lder constant obtained according to the trial results,
\begin{equation}\label{eq:20}
m = 
\begin{cases}
\begin{matrix}
rM, & M>0, \\
1, & M= 0,
\end{matrix}
\end{cases}
\end{equation}
is improved estimation of the H{\"o}lder constant due to method's parameter $r>1$, $\sigma$ is some integer parameter that influences the speed of convergence to the global minimum.

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{fig1.png} 
\caption{Contour plots of 2-dimensional convolution.}\label{fig1} 
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig2.png} 
\caption{Contour plots of 2-dimensional convolution.}\label{fig2} 
\end{figure}

@@

\begin{table}
\tbl{Results of the experiment with bi-criteria MCO problems of dimension 2.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & OMOPSO & SMPSO & NSGAII & IBEA & SPEA2 \\ \midrule
  6.62 & 6.44 & 5.93 & 5.88 & 5.71 & 5.65 & 5.61 \\ \bottomrule
\end{tabular}}
\label{tab1}
\end{table}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig3.pdf} 
\caption{HV indices in bi-criteria 2-dimensional case.}\label{fig3} 
\end{figure}

The second experiment was aimed at solving bi-criteria 4-dimensional MCO problems. 10 test problems have been investigated and for each of them, 100 convolutions (\ref{conv_problem}) with uniformly distributed weight coefficients $\lambda$ were minimized by RALR-DA. 
Tab.~\ref{tab2} contains the results achieved by the methods participated in the experiment and Fig.~\ref{fig4} presents them in graphical form.

\begin{table}
\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & SPEA2 & OMOPSO & IBEA & NSGAII & SMPSO   \\ \midrule
  5.04 & 4.71 & 4.53 & 4.45 & 4.39 & 4.39 & 4.37 \\ \bottomrule
\end{tabular}}
\label{tab2}
\end{table}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig4.pdf} 
\caption{HV indices in bi-criteria 4-dimensional case.}\label{fig4} 
\end{figure}

The third collection of test problems consisted of 10 two-dimensional MCO benchmarks with 4 criteria. In this experiment RALR-DA in each MCO problem minimized convolutions (\ref{conv_problem}) corresponding to 100 different values of weight coefficients $\lambda$ uniformly distributed in the domain (\ref{lambda}).
Like  the  previous experiments, the results are presented in Tab.~\ref{tab3} and Fig.~\ref{fig5}.

\begin{table}
\tbl{Results of testing bi-criteria 4-dimensional MCO problems.}
{\begin{tabular}{ccccccc} \toprule
 Pareto & RALR-DA & OMOPSO & SPEA2 & SMPSO & IBEA & NSGAII   \\ \midrule
  23.89 & 23.52 & 22.87 & 22.80 & 22.53 & 20.68 & 19.82 \\ \bottomrule
\end{tabular}}
\label{tab3}
\end{table}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig5.pdf} 
\caption{HV indices in 4-criteria 2-dimensional case.}\label{fig5} 
\end{figure}

As it can be seen from the presented experiments, the proposed method demonstrates the best results among all methods involved in evaluation of their effectiveness based on HV index. Moreover, the efficiency of RALR-DA is close to the best one estimated on a dense uniform grid.

As for metaheuristic algorithms, their results differ in different experiments.  For example, in 4-dimensinal bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria. On the whole, as the most stable metaheuristic method, the algorithm OMOPSO can be recognized.
